### 2016-11-22 thesis/intro/intro.tex
tmp commit

p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
As \citet{Hanahan2011} mentions in their review, ``whereas \gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.

@@

@@

>>>>

p53 protein is another tumour suppressor protein that is closely related to the \gls{rb} protein and its role in cell cycle arrest \citep{Hanahan2011,Levine1997}.
As \citet{Hanahan2011} mentions in their review, ``\gls{rb} transduces growth-inhibitory signals that originate largely outside of the cell, TP53 receives inputs from stress and abnormality sensors that function within the cell's intracellular operating systems''.
The level of expression and its stability is crucial for the function of p53 protein, as the abundance of p53 protein in the cell triggers cell cycle arrest or cell \gls{apoptosis}, depending on the extent of the cell signal \citep{Fridman2003,Hanahan2011,Levine1997}.
When abundant and active, p53 protein transcribes the p21 protein which, through series of events in the p16-cyclin D$_1$-cdk4-\gls{rb} pathway, triggers cell cycle arrest \citep{Levine1997}.

@@

\noindent
Like normal cells, cancer cells also require access to oxygen and nutrient supply from the blood.
However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.

It is thought that the angiogenic ability of the tumour is gained at an early stage of its development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
Like many cellular processes, angiogenesis is controlled by the balance between the inducing and inhibiting agents of angiogenesis.
\Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have inhibitory effect) \citep{Yancopoulos2000}.
Opposing the effect of \Gls{vegfa} is the \gls{tsp1}, a protein that is known to inhibit the process of angiogenesis by affecting the bioavailability of \acrshort{vegf}, as well as promoting apoptosis \citep{Kazerounian2008}.

In general, many cancers stimulate the expression of \gls{vegfa} protein and reduces the expression of \gls{tsp1} \citep{Kazerounian2008}.

@@

\noindent
One of the prime reasons that cancers are so difficult to treat and have high mortality is due to their ability to metastasise to other parts of the body.
Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology to the primary tumour, and thus have to consider multiple treatment plans to cure the patient.

Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site, and eventually metastasise from this site \citep{Hanahan2011,Kalluri2009}.

There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.
